\documentclass[11pt]{article}
\usepackage[letterpaper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newcommand{\TV}{\mathrm{TV}}
\newcommand{\KL}{D_{\mathrm{KL}}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\usepackage{xcolor}

\title{\textbf{The Value of Delayed Information:\\
Predictive-Information Envelopes for Age-Limited Control}\\[4pt]
\large Exact Bayes envelope generally; mutual-information bound universally;\\
closed form in the symmetric Markov case}
\author{Andrew H. Bond,~\emph{Senior Member, IEEE}\\
Department of Computer Engineering, San Jos\'e State University\\
\texttt{andrew.bond@sjsu.edu} ~\textbar~ ORCID: 0009-0003-2599-6158}
\date{Working draft --- v0.1 (theory core complete; battery + related work open)}

\begin{document}
\maketitle

\begin{abstract}
A controller that acts on information of age $D$ --- a sensing delay, a
telemetry lag, a stale cache --- can outperform the best static policy only to
the extent that the delayed observation still \emph{predicts} the current
state. Our companion systems work \cite{infocom} proves this exactly for
symmetric two-state Markov capacity ($\Delta(D)=\tfrac12 r(D)$, the
autocorrelation at the sensing lag) with a $\beta$-mixing converse. This paper
supplies the general theory. We (i) isolate the exact object --- the
\emph{Bayes envelope} $\Delta_Y = \E[\psi(\mu_Y)]-\psi(\bar\mu)$, the expected
excess value of posterior over prior beliefs; (ii) prove the universal
\emph{predictive-information envelope}
$0\le\Delta_Y\le\Omega\sqrt{\tfrac12 I(X;Y)}$ for any bounded objective with
oscillation $\Omega$, by Pinsker's inequality applied to the posterior
process; (iii) derive the \emph{bit-budget corollary}: a telemetry path
delivering $b$ bits per decision caps feedback value at
$\Omega\sqrt{(b\ln 2)/2}$ regardless of what is encoded --- compression can
preserve but never create feedback value; (iv) show the envelope is
\emph{tight to leading order} in the canonical case: for the symmetric binary
chain $I=\tfrac{r^2}{2}+O(r^4)$, so the bound reads
$\Omega(\tfrac r2+O(r^3))$ against the exact $\tfrac12 r$; and (v) prove by a
two-bit construction that no universal equality $\Delta=f(I)$ can hold:
mutual information is a scalar, decision value is a geometry, and two
observations with identical $I$ can carry all or none of the
decision-relevant distinction. A five-domain empirical section
(cluster contention, acoustic, seismic, space weather, bioacoustic)
tests the law and its envelopes on real processes.
\end{abstract}

\section{Introduction}\label{sec:intro}
\todo{motivation prose; position vs.\ AoI (freshness metrics without
decision semantics), value-of-information (Howard \cite{howard1966}; De
Lara--Gossner \cite{delaragossner}), remote estimation, predictive
information (Bialek et al.), delayed-sharing structures (Witsenhausen).
The one-line thesis: \emph{age matters only through predictive
information, and predictive information matters only through the Bayes
envelope.}}

\section{Setting and the exact object}\label{sec:envelope}

Let $X$ be the current exogenous state taking values in a finite set
$\mathcal X$, with marginal $\bar\mu$. A controller observes a delayed
quantity $Y$ (a delayed sample $X_{t-D}$, a delayed history, or any
$\sigma(\mathcal F_{\le t-D})$-measurable statistic), chooses an action $w$,
and receives a bounded one-step reward $R(w,x)$ with oscillation
$\Omega=\max_{w,x}R-\min_{w,x}R$. Define the value of a belief
$\mu\in\Delta(\mathcal X)$:
\[
\psi(\mu)\;=\;\max_{w}\sum_{x}R(w,x)\,\mu(x),
\]
a pointwise maximum of linear functionals, hence convex; and the posterior
$\mu_Y=\mathbb P(X\in\cdot\mid Y)$, with $\E[\mu_Y]=\bar\mu$ by the tower
property. The optimal $Y$-measurable controller attains
$V^\star_Y=\E[\psi(\mu_Y)]$; the best static controller attains
$V_{\mathrm{stat}}=\psi(\bar\mu)$. The exact object of this paper is the
\textbf{Bayes envelope}
\begin{equation}\label{eq:envelope}
\Delta_Y \;=\; V^\star_Y - V_{\mathrm{stat}}
\;=\; \E\bigl[\psi(\mu_Y)\bigr]-\psi(\bar\mu)\;\ge\;0,
\end{equation}
nonnegative by Jensen. Everything that follows is an envelope on
\eqref{eq:envelope}; the companion paper's exact law is the case where
\eqref{eq:envelope} admits a closed form.

\begin{lemma}[Lipschitz value {\cite[App.~A]{infocom}}]\label{lem:lip}
$|\psi(\mu)-\psi(\mu')|\le\Omega\,\|\mu-\mu'\|_{\TV}$, hence
$\Delta_Y\le\Omega\,\E\|\mu_Y-\bar\mu\|_{\TV}$.
\end{lemma}

The total-variation route specializes to the $\beta$-mixing converse of
\cite{infocom} when $Y$ is the full delayed history:
$\Delta(D)\le\Omega\,\beta(D)$.

\section{The predictive-information envelope}\label{sec:mi}

\begin{theorem}[Predictive-information bound]\label{thm:mi}
For any bounded objective with oscillation $\Omega$ and any delayed
observable $Y$,
\begin{equation}\label{eq:mi}
0\;\le\;\Delta_Y\;\le\;\Omega\,\sqrt{\tfrac12\,I(X;Y)}
\qquad\text{($I$ in nats; in bits, }\Omega\sqrt{\tfrac{\ln 2}{2}\,I_{\mathrm{bits}}}\text{).}
\end{equation}
\end{theorem}

\begin{proof}
By Lemma~\ref{lem:lip}, $\Delta_Y\le\Omega\,\E\|\mu_Y-\bar\mu\|_{\TV}$.
Pinsker's inequality gives, for every realization $y$,
$\|\mu_y-\bar\mu\|_{\TV}\le\sqrt{\tfrac12\KL(\mu_y\|\bar\mu)}$. Taking the
expectation over $Y$ and applying Jensen to the concave square root,
\[
\E\|\mu_Y-\bar\mu\|_{\TV}
\;\le\;\E\sqrt{\tfrac12\KL(\mu_Y\|\bar\mu)}
\;\le\;\sqrt{\tfrac12\,\E\,\KL(\mu_Y\|\bar\mu)}
\;=\;\sqrt{\tfrac12\,I(X;Y)},
\]
the last equality being the definition of mutual information as the expected
KL divergence of posterior from prior.
\end{proof}

\begin{corollary}[Bit-limited telemetry]\label{cor:bits}
If the sensing path delivers at most $b$ bits per decision --- $Y=g(Z)$ with
$Z$ the raw delayed signal and $H(Y)\le b\ln 2$ nats --- then
$\Delta_Y\le\Omega\sqrt{(b\ln 2)/2}$, and moreover
$\Delta_Y\le\Omega\sqrt{\tfrac12 I(X;Z)}$ regardless of $g$ (data
processing). Compression can \emph{preserve} feedback value; no encoding can
create it.
\end{corollary}

\begin{proposition}[Leading-order tightness in the canonical case]
\label{prop:tight}
For the symmetric two-state Markov chain with lag-$D$ autocorrelation
$r=r(D)$,
\[
I(X_t;X_{t-D})=\tfrac12\bigl[(1{+}r)\ln(1{+}r)+(1{-}r)\ln(1{-}r)\bigr]
=\tfrac{r^2}{2}+\tfrac{r^4}{12}+O(r^6),
\]
so the right side of \eqref{eq:mi} equals
$\Omega\bigl(\tfrac r2+O(r^3)\bigr)$, matching the exact law
$\Delta(D)=\tfrac12 r(D)$ of \cite{infocom} to leading order (at the
objective normalization $\Omega=1$). The information envelope is not loose
where it matters.
\end{proposition}

\begin{proof}
The joint distribution places mass $\tfrac{1+r}{4}$ on each agreeing pair and
$\tfrac{1-r}{4}$ on each disagreeing pair; the stated $I$ follows by direct
computation, and the expansion from
$(1\pm r)\ln(1\pm r)=\pm r+\tfrac{r^2}{2}\mp\tfrac{r^3}{6}+\cdots$. Then
$\sqrt{\tfrac12 I}=\tfrac r2\sqrt{1+\tfrac{r^2}{6}+O(r^4)}$.
\end{proof}

\section{No universal equality \texorpdfstring{$\Delta=f(I)$}{Delta=f(I)}}
\label{sec:noequality}

\begin{proposition}[Information is a scalar; value is a geometry]
\label{prop:noneq}
There is a state space, a bounded objective, and two observations
$Y_1, Y_2$ with $I(X;Y_1)=I(X;Y_2)>0$ but
$\Delta_{Y_1}=\tfrac{\Omega}{2}\cdot\tfrac12>0=\Delta_{Y_2}$. Hence no
function $f$ with $\Delta_Y=f(I(X;Y))$ exists for all experiments.
\end{proposition}

\begin{proof}
Let $X=(B_1,B_2)$ be two independent fair bits and let the objective depend
only on $B_1$: actions $w\in\{0,1\}$, $R(w,x)=\Omega\,\mathbf 1[w=B_1]$
(so the oscillation is $\Omega$). Static value: $\Omega/2$. Let $Y_1=B_1$
and $Y_2=B_2$. Both have $I(X;Y_i)=\ln 2$. Observing $Y_1$ yields
$V^\star=\Omega$, so $\Delta_{Y_1}=\Omega/2$; observing $Y_2$ leaves the
posterior over $B_1$ uniform, so $\Delta_{Y_2}=0$. The two experiments share
every information quantity in $I(X;Y)$ yet differ maximally in value: the
envelope \eqref{eq:mi} is attained within a constant on one and vacuous on
the other.
\end{proof}

\begin{remark}
This is the precise sense in which the companion paper's binary law is
special: there the state is one decision-relevant bit, so the geometry
collapses to a one-parameter family and the envelope inverts to an exact
closed form. The general exact object remains \eqref{eq:envelope}.
\end{remark}

\section{Envelope comparison}\label{sec:comparison}

The three envelopes are not interchangeable, and their relationship is an
ordered pair of relaxations of one master quantity.

\begin{proposition}[Master envelope and universal ordering]\label{prop:order}
For any delayed observable $Y$ measurable w.r.t.\ the delayed history
$\mathcal F_{\le t-D}$,
\[
\Delta_Y
\;\le\;\Omega\,\E\|\mu_Y-\bar\mu\|_{\TV}
\;\le\;\Omega\,\min\Bigl\{\beta(D),\ \sqrt{\tfrac12 I(X;Y)}\Bigr\},
\]
where $\beta(D)$ is the $\beta$-mixing coefficient of the state process at
lag $D$. The expected posterior movement
$\E\|\mu_Y-\bar\mu\|_{\TV}$ is the master envelope; the mixing and
information envelopes are its two relaxations.
\end{proposition}

\begin{proof}
The first inequality is Lemma~\ref{lem:lip}. For the second: since
$Y$ is $\mathcal F_{\le t-D}$-measurable and $\sigma(X)\subseteq
\mathcal F_{\ge t}$, conditioning on the coarser $Y$ can only shrink
expected posterior movement, and
$\E\|\mu_{\mathcal F_{\le t-D}}-\bar\mu\|_{\TV}\le\beta(D)$ because the
$\beta$-coefficient takes its supremum over the larger future field
\cite[App.~A]{infocom}. The information relaxation is the Pinsker step in the
proof of Theorem~\ref{thm:mi}.
\end{proof}

\begin{proposition}[The two relaxations are incomparable]\label{prop:incomp}
Neither relaxation dominates the other.
\begin{enumerate}
\item[(i)] \emph{Mixing strictly tighter:} for the symmetric two-state chain
with $Y=X_{t-D}$, $\beta(D)=\tfrac12 r$ while
$\sqrt{\tfrac12 I}=\tfrac r2\sqrt{1+\tfrac{r^2}{6}+O(r^4)}>\beta(D)$ for
every $r\in(0,1)$.
\item[(ii)] \emph{Information arbitrarily tighter:} for every $K\ge 2$ there
is a stationary state process and objective with
$\beta(D)=\tfrac12$ for all $D$ (the mixing envelope never decays) while
$\sqrt{\tfrac12 I(X;\mathcal F_{\le t-D})}=\sqrt{\ln 2/(2K)}\to 0$, and
indeed $\Delta=\Omega/(2K)$.
\end{enumerate}
\end{proposition}

\begin{proof}
(i) is Proposition~\ref{prop:tight} together with
$\beta(D)=\tfrac12 r(D)$ for the binary chain \cite[Cor.~binary]{infocom}.
(ii) \emph{The preview process.} Let $c_t$ be i.i.d.\ fair bits, let
$\Theta\sim\mathrm{Unif}\{0,\dots,K{-}1\}$ be an independent phase, and let
the observable state be $X_t=(c_t,p_t)$ with
\[
p_t \;=\; \begin{cases} c_{t+m} & t\equiv\Theta \pmod K\\
\bot & \text{otherwise},\end{cases}
\qquad m = D+K,
\]
a sensor that occasionally publishes a \emph{forecast} $m$ steps ahead. The
objective depends only on the current payoff bit:
$R(w,x)=\Omega\,\mathbf 1[w=c]$. \emph{Mixing:} the window
$[t-m,\,t-D]$ has length $K$, hence contains exactly one preview slot
$s\equiv\Theta$; that preview reveals $c_{s+m}\in\{c_t,\dots,c_{t+K}\}$, a
present-or-future payoff bit, so some event in $\mathcal F_{\ge t}$ has
conditional probability in $\{0,1\}$ against marginal $\tfrac12$:
$\beta(D)=\tfrac12$ for every $D$. \emph{Information:} the delayed history
determines $c_t$ exactly when $t-m\equiv\Theta\pmod K$ (probability $1/K$
over the phase) and is independent of $c_t$ otherwise, so
$I(c_t;\mathcal F_{\le t-D})=\tfrac{\ln 2}{K}$. \emph{Value:} feedback wins
$\Omega/2$ on the revealed fraction $1/K$ and nothing elsewhere:
$\Delta=\Omega/(2K)$. The mixing envelope is vacuous; the information
envelope $\Omega\sqrt{\ln 2/(2K)}$ decays with $K$ (at rate $K^{-1/2}$
against the true $K^{-1}$ --- loose, but not blind).
\end{proof}

\begin{remark}[Why keep the information form at all]
Proposition~\ref{prop:order} shows the TV master envelope is always at least
as tight, and (i) shows mixing can beat information. Three properties are the
information envelope's own: it is \emph{estimable} --- the plug-in
$\hat I(X_t;X_{t-D})$ in \S\ref{sec:battery} needs only pair counts, while
the $\beta$-coefficient's supremum over the future $\sigma$-field is not
estimable without a process model; it \emph{composes with channels} --- data
processing gives Corollary~\ref{cor:bits}, and no analogous statement exists
for $\beta$ (mixing has no channel semantics); and it \emph{adds across
sensors} via the chain rule, giving budget statements for multi-sensor
telemetry. In short: mixing is the sharper analysis tool when a model is in
hand; information is the operational one when only data or a bit budget is.
\end{remark}

\begin{table}[h]\centering\small
\begin{tabular}{llll}
\toprule
Envelope & Bound on $\Delta_Y$ & Needs & Blind spot \\
\midrule
master (TV) & $\Omega\,\E\|\mu_Y-\bar\mu\|_{\TV}$ & posterior process & not estimable directly \\
mixing & $\Omega\,\beta(D)$ & process model & future events beyond $X_t$ (Prop.~\ref{prop:incomp}ii) \\
information & $\Omega\sqrt{I(X;Y)/2}$ & pair samples & decision geometry (Prop.~\ref{prop:noneq}) \\
bit budget & $\Omega\sqrt{(b\ln 2)/2}$ & channel rate only & everything but the rate \\
\bottomrule
\end{tabular}
\caption{The envelope family. Each successive row needs less knowledge and
pays for it in tightness; the battery of \S\ref{sec:battery} measures the
first three on real processes.}
\label{tab:envelopes}
\end{table}

\section{Exactness beyond two states}\label{sec:multistate}
\todo{layer-cake lift (cite \cite[Prop.~2]{infocom}): exact
$\Delta(D)=\sum_u\pi_u r_u(D)$ for monotone-threshold objectives ---
already a multi-state exact law; characterize the objective class for which
the Bayes envelope admits closed form; asymmetric two-state case.}

\section{The law and its envelopes across domains}\label{sec:battery}

The companion paper had room for one paragraph of this study; the full
protocol and results are reported here. The question has two halves. First,
\emph{the law}: is $\Delta(D)=\tfrac12 r(D)$ --- exact for symmetric binary
Markov capacity --- also the operative value of feedback on \emph{real}
binary state processes, i.e., does history beyond the single delayed sample
add value? Second, \emph{the envelope}: does the measured value respect, and
how tightly does it approach, the predictive-information bound of
Theorem~\ref{thm:mi}?

\subsection{Protocol}

\textbf{States.} Binary regime series from four physical domains plus
controls: acoustic source azimuth and voice activity (LOCATA, 120\,Hz);
sperm-whale coda type (Dominica corpus, per-exchange); GOES space-weather
series (X-ray flux, magnetometer, proton flux; 1--5-minute cadence); seismic
station regime (ANMO, 1-minute); a synthetic two-state Markov chain
(flip probability $0.02$) as positive reference; an i.i.d.\ fair-bit series
as null. Each series is binarized by the domain's registered thresholding
(the companion study's preparation scripts, unchanged).

\textbf{The two estimands.} For each lag $D$ on a per-domain grid:
(i) the pooled lag-$D$ autocorrelation $r(D)$, giving the law value
$\tfrac12 r(D)$; (ii) the value of the best \emph{delayed-history} policy
$\widehat\Delta_{\mathrm{hist}}(D)$: a $k{=}6$-lag lookup predictor trained
and evaluated by blocked cross-validation (5 folds, contiguous blocks, no
leakage across fold boundaries), reported as accuracy above chance ---
the operational value of feedback for the $0/1$ prediction objective
($\Omega=1$, for which the binary law reads exactly $\tfrac12 r$).

\textbf{Surrogate calibration.} The null for ``history adds nothing beyond
the law'' is a population of $60$ Markov surrogates per domain, each matched
to the real series' lag-1 transition probabilities and evaluated by the
identical pipeline. The gain
$\widehat\Delta_{\mathrm{hist}}-\tfrac12 r$ is called significant only if
its maximum over the lag grid exceeds the surrogate maxima at $p<0.05$
\emph{and} clears a practical floor of $0.03$ (twice the i.i.d.\ null's
estimator fluctuation) --- the dual criterion of the companion study.

\textbf{The envelope measurement.} New here: for each domain and lag, the
plug-in mutual information $\widehat I(X_t;X_{t-D})$ from pooled pair counts
($2{\times}2$ table; plug-in bias $\le 3/(2N)$ nats at the smallest pooled
$N$ in the battery, negligible at the reported scale), and the envelope
$\sqrt{\widehat I/2}$ of Theorem~\ref{thm:mi}. Because
$\widehat\Delta_{\mathrm{hist}}$ is a noisy estimate, the registered check is
\[
\widehat\Delta_{\mathrm{hist}}(D)\;\le\;\sqrt{\widehat I(D)/2}\;+\;
\varepsilon_{\mathrm{est}},
\]
with $\varepsilon_{\mathrm{est}}$ the maximum absolute fluctuation of the
same estimator on the i.i.d.\ null --- the null calibrates the noise floor,
so an apparent ``violation'' smaller than what pure noise produces on a
zero-information series is reported as estimator noise, not physics.

\subsection{Results}

Table~\ref{tab:battery} reports, per domain: the maximum history gain beyond
the law with its surrogate $p$-value; the per-domain noise bound
$\varepsilon_{\mathrm{dom}}$ (the maximum of the surrogates' $97.5\%$ gain
band --- surrogates have zero true history gain by construction and share the
domain's series lengths, so they calibrate its estimator noise); the maximum
excess of the history policy over the \emph{single-lag} envelope
$\sqrt{\hat I(X_t;X_{t-D})/2}$; and the tightness ratio
envelope$/$law (median over lags with $r>0.05$).

\begin{table}[h]\centering\small
\begin{tabular}{lrrrrrl}
\toprule
Domain & gain$_{\max}$ & $p$ & $\varepsilon_{\mathrm{dom}}$ &
$\max(\widehat\Delta-\mathrm{env}_1)$ & env$/$law & verdict \\
\midrule
Two-state Markov (ref)   & $0.030$ & $.38$ & $0.069$ & $+0.023$ & $1.001$ & law exact \\
i.i.d.\ null (control)   & $0.012$ & $.68$ & $0.016$ & $+0.012$ & ---     & $\Delta\approx 0$ \\
LOCATA azimuth           & $0.008$ & $.05$ & $0.007$ & $-0.013$ & $1.099$ & law exact \\
Sperm-whale codas        & $0.046$ & $.98$ & $0.112$ & $+0.045$ & $1.010$ & law exact \\
GOES protons             & $0.029$ & $.77$ & $0.048$ & $+0.030$ & $0.998$ & law exact \\
\addlinespace
LOCATA VAD               & $0.100$ & $.00$ & $0.059$ & $\mathbf{+0.128}$ & $0.940$ & history adds \\
GOES X-ray flux          & $0.185$ & $.00$ & $0.100$ & $+0.071$ & $0.937$ & history adds \\
GOES magnetometer        & $0.147$ & $.00$ & $0.070$ & $+0.041$ & $0.972$ & history adds \\
Seismic ANMO             & $0.367$ & $.00$ & $0.037$ & $\mathbf{+0.157}$ & $0.999$ & history adds \\
\bottomrule
\end{tabular}
\caption{Battery v3 (nine series, four physical domains + controls;
\texttt{battery\_v3.json}). Upper block: the law is the whole value of
feedback and the single-lag envelope is respected within noise. Lower block:
the surrogate test certifies history value beyond the law
($p{<}.05$ and gain above the $0.03$ floor), and exactly there the history
policy can exceed the \emph{single-sample} envelope --- bold where the
excess clears $\varepsilon_{\mathrm{dom}}$, certifying window information
beyond the single lag (see text).}
\label{tab:battery}
\end{table}

Three readings, in increasing order of consequence.

\textbf{The law, confirmed where it should hold.} On the Markov reference
and the three domains whose surrogate test finds nothing beyond the law
(azimuth, whale codas, protons), the best delayed-history policy sits on
$\tfrac12 r(D)$ across the lag grid --- the single delayed sample is the
whole value of feedback, in acoustics, bioacoustics, and space weather alike.

\textbf{Tightness, observed in the wild.} The envelope$/$law column is
Proposition~\ref{prop:tight} measured: $1.001$ on the reference chain
(small $r$: the envelope sits \emph{on} the law), rising to $1.099$ on
azimuth exactly as the $\sqrt{1+r^2/6}$ correction predicts for its
slowly-decaying (large-$r$) regime. The information envelope is not a loose
theoretical courtesy; on Markov-regime data it \emph{is} the law, to within
a predicted $O(r^2)$ factor.

\textbf{The envelope as a non-Markovianity witness.} In the lower block the
history policy exceeds the single-lag envelope --- decisively on VAD
($+0.128$ against a noise bound of $0.059$) and seismic regime
($+0.157$ vs.\ $0.037$). This is not a counterexample to
Theorem~\ref{thm:mi}: the history policy's observable is the $6$-lag
window $Y$, and the theorem bounds it by $\sqrt{I(X_t;Y)/2}$ for
\emph{that} $Y$. Read backwards, the theorem converts the measured
exceedance into a \emph{certified information lower bound}:
\[
I\bigl(X_t;\,\text{window}\bigr)\;\ge\;2\,\widehat\Delta_{\mathrm{hist}}^2,
\]
which on the seismic series certifies $\ge 0.072$ nats of window information
at a lag where the single sample carries $0.002$ --- a $30\times$
concentration of predictive information in history beyond the last sample
(aftershock clustering, in information units). The two blocks of
Table~\ref{tab:battery} are therefore not ``pass'' and ``fail'': they are
the theorem operating in both directions --- as an envelope on Markov-regime
processes, and as a quantitative witness of history structure on
non-Markov ones. The dual use is registered in the companion study's terms:
the same surrogate machinery that certifies the law's exactness upper-block
certifies the window-information excess lower-block.

\section{Related work}\label{sec:related}
\todo{Howard VoI \cite{howard1966}; De Lara--Gossner \cite{delaragossner};
AoI surveys; remote estimation / Witsenhausen delayed sharing; Bialek
predictive information; rate-distortion with side information; NCS.}

\section{Conclusion}\label{sec:conclusion}
\todo{}

\begin{thebibliography}{9}
\bibitem{infocom}
A.~H.~Bond. Polite bursting: the value of delayed feedback in age-limited
admission control. Under submission, IEEE INFOCOM 2027.

\bibitem{howard1966}
R.~A.~Howard. Information value theory.
\emph{IEEE Trans.\ Systems Science and Cybernetics} 2(1):22--26, 1966.

\bibitem{delaragossner}
M.~De~Lara, O.~Gossner. Payoffs-beliefs duality and the value of
information. \emph{SIAM J.\ Optimization} 30(1):287--310, 2020.
\end{thebibliography}

\end{document}
