% nats-bursting -- INFOCOM model, STRENGTHENED (drop-in replacement for the
% model/controller sections of infocom_model.tex). The thermal micro-tier moves
% to the CANOPIE-HPC companion; this file is single-tier (admission) and adds the
% Age-of-Information feedback-usefulness limit that turns the Poisson null into
% the paper's central result.
%
% Preamble (ensure present in the main file):
%   \usepackage{amsmath,amssymb,amsthm}
%   \newtheorem{theorem}{Theorem}
%   \newtheorem{lemma}{Lemma}
%   \newtheorem{corollary}{Corollary}
%   \newtheorem{proposition}{Proposition}
%   \theoremstyle{definition}\newtheorem{assumption}{Assumption}

\section{System Model: Polite Bursting under Age-Limited Sensing}
\label{sec:model}

A \emph{guest} bursts batch GPU pods onto a shared cluster and is bound by a
fair-use policy: at most $K$ concurrent pods, and each admitted pod must sustain
a utilization floor (else it is a poor citizen, risking throttling or a ban).
Time is slotted into control epochs $t\in\mathbb{N}$. Let $a_t\in\{0,\dots,C\}$ be
the capacity the cluster can \emph{productively} give the guest at epoch $t$; it
is set by the other tenants---exogenous, stochastic, and \textbf{not directly
observable}. The guest chooses a window $w_t\in\{0,\dots,K\}$. Useful goodput is
$G_t=\min(w_t,a_t)$; the surplus $(w_t-a_t)^{+}$ is \emph{over-admitted}: those
pods sit below the utilization floor and mark the guest impolite. With
$\rho\triangleq\limsup_T\frac1T\sum_{t<T}\mathbf 1[w_t>a_t]$ the long-run
over-admission fraction, the guest solves
\begin{equation}\label{eq:polite}
\max_{\{w_t\}}\ \mathbb{E}[\min(w_t,a_t)]
\quad\text{s.t.}\quad
\underbrace{w_t\le K\ \forall t}_{\text{hard cap}},\qquad
\underbrace{\rho\le\epsilon}_{\text{policy}} .
\end{equation}
The constraint is \emph{two-sided}---stay below the cap yet above the floor---which
already separates it from one-sided (loss-only) congestion control.

\paragraph{Age-limited sensing.}
The guest does not observe $a_t$; it \emph{probes} cluster state, and the probe
is stale. We model this as an \emph{age of information} $D$: at epoch $t$ the
controller's freshest reading of capacity has age $D$, i.e.\ admissible policies
are causal functions of $\{a_{s-D}\}_{s\le t}$. $D$ is a physical, \emph{measured}
quantity (Sec.~\ref{sec:eval}: $D\!\approx\!2.5$\,s, dominated by CUDA cold-start,
not the $\sim\!27$\,ms poll). Age-limited sensing is the crux: a controller can
only act on a $D$-old picture of a world the neighbours are still changing.

\paragraph{The contended unit.}
Politeness is decided at the margin. Fix attention on the single contended
resource unit whose availability fluctuates with the neighbour, and let
$x_t\in\{0,1\}$ indicate that it is \emph{free} ($x_t=1$) or \emph{taken}
($x_t=0$); admitting it when taken is the over-admission the policy punishes. We
take $x_t$ stationary with $\Pr[x_t=1]=\tfrac12$ (a fully contended unit; the
asymmetric case only rescales the constants below). The guest's marginal decision
$m_t\in\{0,1\}$ is a (possibly randomized) causal function of the age-$D$
observation $\hat x_t\triangleq x_{t-D}$. Marginal goodput and over-admission are
\begin{equation}\label{eq:margin}
g \triangleq \Pr[m_t=1,\,x_t=1],\qquad
\rho \triangleq \Pr[m_t=1,\,x_t=0].
\end{equation}
Everything below concerns the achievable $(\rho,g)$ region under age $D$.

\begin{proposition}[Hard-cap safety]\label{prop:safety}
Any window update that clamps additive increase at $K$ and never increases $w_t$
on a congestion signal keeps $w_t\le K$ for all $t$ deterministically; the pod-cap
is never violated.
\end{proposition}
\begin{proof}
Induction on $t$: increase is clamped to $K$, decrease cannot exceed the current
value.
\end{proof}

\section{The Feedback-Usefulness Limit}
\label{sec:limit}

We now characterize \emph{exactly} how much closed-loop admission can gain over
open-loop (static) admission when the only signal is a $D$-old observation. The
answer is a single, interpretable quantity: the autocorrelation of capacity at the
sensing lag.

\begin{lemma}[Predictive disagreement]\label{lem:q}
Let $q(D)\triangleq\Pr[x_t\neq x_{t-D}]$ be the lag-$D$ disagreement probability
and $r(D)\triangleq\mathbb{E}[(2x_t\!-\!1)(2x_{t-D}\!-\!1)]$ the lag-$D$
autocorrelation of the centered availability. Then $r(D)=1-2q(D)\in[-1,1]$, and by
stationarity and symmetry the joint law of $(\hat x_t,x_t)$ is
$\Pr[\hat x_t{=}i,x_t{=}j]=\tfrac14\big(1+(2i{-}1)(2j{-}1)\,r(D)\big)$.
For a symmetric two-state Markov capacity with per-epoch flip probability
$p\in(0,\tfrac12)$, $r(D)=(1-2p)^{D}=e^{-D/\tau_c}$ with correlation time
$\tau_c\triangleq-1/\ln(1-2p)\approx 1/(2p)$.
\end{lemma}
\begin{proof}
$r=\mathbb{E}[(2x_t-1)(2x_{t-D}-1)]=\Pr[\text{agree}]-\Pr[\text{disagree}]=1-2q$.
The joint follows from the two symmetric marginals and the single correlation.
For the Markov chain the $D$-step transition is
$P^D=\tfrac12\!\left(\begin{smallmatrix}1+\lambda^D&1-\lambda^D\\1-\lambda^D&1+\lambda^D\end{smallmatrix}\right)$
with $\lambda=1-2p$, so $\Pr[\text{disagree}]=\tfrac12(1-\lambda^D)$ and
$r=\lambda^D$.
\end{proof}

\begin{theorem}[Achievable region and the feedback advantage]\label{thm:limit}
Under age-$D$ sensing the achievable $(\rho,g)$ region of
\eqref{eq:margin} is the convex hull of
$\{(0,0),\,(\tfrac{q}{2},\tfrac{1-q}{2}),\,(\tfrac12,\tfrac12)\}$, with $q=q(D)$.
The best \emph{static} (observation-oblivious) policy is confined to the diagonal
$g=\rho$. Consequently the goodput gain of feedback over static, maximized over
matched over-admission levels, is
\begin{equation}\label{eq:adv}
\boxed{\;\Delta(D)\;=\;\tfrac12\,r(D)\;=\;\tfrac12\big(1-2q(D)\big)\;}
\end{equation}
attained at the kink $\rho=q/2$ by the Bayes threshold policy $m_t=\hat x_t$ (at a
lower level $\rho<q/2$ the gain is the smaller $\rho(1-2q)/q$); i.e.\ \emph{the peak
value of closed-loop admission is exactly half the autocorrelation of available
capacity at the sensing lag}.
\end{theorem}
\begin{proof}
A policy is a pair $(\theta_1,\theta_0)$, $\theta_i=\Pr[m_t{=}1\mid\hat x_t{=}i]$.
Using Lemma~\ref{lem:q},
$g=\tfrac12\big(\theta_1(1{-}q)+\theta_0 q\big)$ and
$\rho=\tfrac12\big(\theta_1 q+\theta_0(1{-}q)\big)$.
This linear map sends the square $[0,1]^2$ to the quadrilateral with vertices at
$(\theta_1,\theta_0)\in\{(0,0),(1,0),(1,1),(0,1)\}$, i.e.\
$(0,0),(\tfrac q2,\tfrac{1-q}2),(\tfrac12,\tfrac12),(\tfrac{1-q}2,\tfrac q2)$; its
upper boundary is the stated hull. Maximizing $g-\lambda\rho$ turns on $\theta_1$
before $\theta_0$ because $q<\tfrac12$ makes the free-observation state strictly
more goodput-efficient, so the max-goodput point for $\rho\le q/2$ is the kink
$(\tfrac q2,\tfrac{1-q}2)$, achieved by $m_t=\hat x_t$. A static policy ignores
$\hat x_t$ ($\theta_1=\theta_0=\theta$), giving $g=\rho=\theta/2$: the diagonal.
The vertical gap at $\rho=q/2$ is $\tfrac{1-q}{2}-\tfrac q2=\tfrac{1-2q}{2}=\tfrac12 r(D)$.
\end{proof}

\begin{corollary}[Impossibility as sensing ages]\label{cor:imp}
$\Delta(D)\!\to\!0$ as $r(D)\!\to\!0$. For any capacity process that is mixing
($r(D)\!\to\!0$), and in particular for the Markov model
$\Delta(D)=\tfrac12 e^{-D/\tau_c}$, no causal controller with sensing age $D$ can
outperform the static budget once $D\gg\tau_c$: closed-loop admission is useful
\emph{iff} the guest can sense the neighbour faster than the neighbour changes.
At $D=0$, $\Delta=\tfrac12$ (capacity is perfectly trackable); the advantage halves
every $\tau_c\ln 2$ epochs of added age. App.~\ref{app:beta} extends this
impossibility to arbitrary (non-binary, non-Markov) capacity processes, with the
decay rate set by the process's $\beta$-mixing coefficient.
\end{corollary}

\begin{corollary}[The square/Poisson experiments are two points on \eqref{eq:adv}]
\label{cor:exp}
The contention study of Sec.~\ref{sec:eval} instantiates exactly this two-state
$C{=}2$ model. \emph{Structured (square)} load has $\tau_c\gg D$, so
$r(D)\!\approx\!1$ and feedback is near-maximally useful (low $\rho$, large goodput
gain); \emph{memoryless (Poisson)} load drives $\tau_c\lesssim D$, so
$r(D)\!\approx\!0$ and feedback collapses onto static---the pre-registered null.
The single curve \eqref{eq:adv} predicts both regimes and the crossover between
them.
\end{corollary}

The binary reduction is not a special case but the atom of the general problem.

\begin{proposition}[Layer-cake reduction for $C>2$]\label{prop:layercake}
Let $a_t\in\{0,\dots,C\}$ and score over-admission by the surplus \emph{amount}
$\bar\rho\triangleq\mathbb{E}[(w_t-a_t)^{+}]$. With $x^u_t\triangleq\mathbf1[a_t\ge u]$
and $\pi_u\triangleq\Pr[x^u_t=1]$, both objectives decompose additively over units,
$\min(w_t,a_t)=\sum_{u\ge1}\mathbf1[w_t\ge u]\,x^u_t$ and
$(w_t-a_t)^{+}=\sum_{u\ge1}\mathbf1[w_t\ge u](1-x^u_t)$. The age-$D$ optimum is the
delayed-capacity window $w_t=a_{t-D}$, and its total goodput advantage over the
best comonotone static budget at matched surplus is
\begin{equation}\label{eq:adv-multi}
\Delta(D)=\sum_{u=1}^{K}\pi_u\,r_u(D),\qquad r_u(D)=\mathrm{corr}\big(x^u_t,x^u_{t-D}\big),
\end{equation}
which for a single symmetric contended unit ($\pi_u=\tfrac12$) recovers
$\Delta=\tfrac12 r(D)$.
\end{proposition}
\begin{proof}
The two identities are the layer-cake (Fubini) expansions of $\min$ and $(\cdot)^+$.
Setting $w_t=a_{t-D}$ makes $\mathbf1[w_t\ge u]=x^u_{t-D}=\hat x^u_t$, i.e.\ the
Thm.~\ref{thm:limit} threshold policy applied independently to each layer; the
layers' nesting $x^1_t\ge x^2_t\ge\cdots$ is inherited by the observations, so the
per-unit thresholds compose into a valid scalar window. For an asymmetric layer
with $\Pr[x^u{=}1]=\pi_u$, the same LP as Thm.~\ref{thm:limit} gives per-unit
advantage $\mathrm{Cov}(x^u_t,x^u_{t-D})/(1-\pi_u)=\pi_u r_u(D)$ over the static
policy matching its marginal; additivity of goodput and surplus sums these to
\eqref{eq:adv-multi}.
\end{proof}

At $C{=}2$ the contended layer alone fluctuates ($\pi_1{=}1,r_1{=}0$;
$\pi_2{=}\tfrac12$), so \eqref{eq:adv-multi} collapses to $\tfrac12 r_2(D)$: the
binary law is exactly the $C{=}2$ instance, and the experiments measure it directly.

\section{Achievability: A Regime-Adaptive Controller}
\label{sec:controller}

Theorem~\ref{thm:limit} exposes a gap that a \emph{fixed} controller cannot close:
additive-increase/multiplicative-decrease (AIMD) tracks a slowly varying neighbour
and approaches the kink $(\tfrac q2,\tfrac{1-q}2)$ when $\tau_c\gg D$, but
\emph{over-admits} when $\tau_c\lesssim D$ (it keeps probing a signal that is pure
noise); a static budget is safe everywhere but forfeits $\Delta(D)$ when
$\tau_c\gg D$. The optimal policy must know which regime it is in. Since $r(D)$ is
itself \emph{measurable from the probe stream}, the guest can estimate it online
and switch.

\begin{center}\small
\fbox{\parbox{0.94\columnwidth}{
\textbf{Regime-adaptive admission.} Maintain the empirical lag-$D$ correlation
$\hat r_n$ from the last $n$ probes. Each epoch:
(i) if $\hat r_n>\gamma$ (trackable): run AIMD with the safe $(\alpha,\beta)$ of
Prop.~\ref{prop:bound}, converging toward the Bayes kink;
(ii) else (untrackable): fall back to the static budget $w_t\equiv\lfloor \bar a\rfloor$.
The threshold $\gamma$ is set from the target $\epsilon$ via \eqref{eq:adv}.}}
\end{center}

\begin{proposition}[Sawtooth realization of the kink]\label{prop:bound}
Under stationary capacity $a$ and (A1)--(A3) of the tracking regime, AIMD forms a
periodic sawtooth about $a$ with long-run over-admission
$\rho=D\alpha/\big((1-\beta)(a+D\alpha)\big)\approx D\alpha/((1-\beta)a)$; choosing
$\alpha\le\epsilon(1-\beta)a/D$ gives $\rho\le\epsilon$, and $\rho\to0$ as $D\to0$.
\end{proposition}
\begin{proof}[Proof sketch]
After a backoff the window restarts at $\beta(a+D\alpha)$ and rises by $\alpha$ per
clean epoch, overshooting $a$ for $D$ epochs before detection triggers the next
backoff at peak $a+D\alpha$; the fraction of violating epochs per cycle gives
$\rho$. (This is the dynamic approach to the static-capacity kink of
Thm.~\ref{thm:limit}; $D$ enters both here and in \eqref{eq:adv}, the
systems$\leftrightarrow$theory hinge the paper turns on.)
\end{proof}

\begin{theorem}[Uniform near-optimality]\label{thm:adapt}
Let capacity be a symmetric two-state process with correlation time $\tau_c$
unknown to the guest. The empirical correlation is consistent,
$\hat r_n\!\to\!r(D)$ a.s.\ (ergodicity), with
$\Pr[|\hat r_n-r(D)|>t]\le 2e^{-cnt^2}$ for a mixing constant $c>0$
(Hoeffding for functions of a geometrically mixing chain). Hence the
regime-adaptive controller attains goodput within
$\Delta(D)-O\!\big(\sqrt{\log n/n}\big)$ of the age-$D$ optimum
\emph{uniformly in} $\tau_c$, whereas (i) fixed AIMD incurs excess over-admission
$\Omega(1)$ as $\tau_c/D\!\to\!0$, and (ii) the fixed static budget forfeits
goodput $\Omega(1)$ as $\tau_c/D\!\to\!\infty$. The regret rate is
$O(D\sqrt{\log N/(\gamma N)})$ over horizon $N$, with $\gamma$ the chain's spectral
gap (App.~\ref{app:regret}).
\end{theorem}
\begin{proof}[Proof sketch]
Consistency and the concentration bound give a correct regime decision after
$n=O(\log(1/\delta)/(\,r{-}\gamma\,)^2)$ probes w.p.\ $1-\delta$; conditioned on the
correct branch, Thm.~\ref{thm:limit} (static branch) and Prop.~\ref{prop:bound}
(AIMD branch) bound the per-epoch gap, and averaging over the $O(1)$ misclassified
prefix yields the stated regret. The separations (i)--(ii) are the two endpoints of
Cor.~\ref{cor:imp}: at small $\tau_c$ AIMD sits at $\rho\!\approx\!\tfrac14$ while
the optimum is $\rho\!\to\!0$; at large $\tau_c$ static forfeits
$\Delta\!\to\!\tfrac12$.
\end{proof}

\paragraph{Scope, honestly.}
Three assumptions earn the closed form. (i) \emph{Binary/symmetric per layer:} the
law \eqref{eq:adv} holds for any symmetric stationary $x_t$ via $r(D)$---not only
Markov---and Prop.~\ref{prop:layercake} lifts it to any integer $C$ once
over-admission is scored by \emph{amount} rather than by the $0/1$ indicator (the
two coincide at $C{=}2$, so our testbed is unaffected). (ii) \emph{Stationarity:}
$r(D)$ is a lag-$D$ autocorrelation of a stationary process; under slow drift it is
replaced by a local estimate, which is exactly what the regime-adaptive controller
tracks (\S\ref{sec:controller}). (iii) \emph{Fixed age $D$:} if the freshest
sample has a random age $A$ whose value the controller can read, the advantage is
the age-averaged $\Delta=\tfrac12\,\mathbb{E}[r(A)]$---the same law under the age
law---and the impossibility $\Delta\to0$ persists whenever $A\gg\tau_c$ almost
surely. A matching
lower bound in the process's $\beta$-mixing coefficient, for general non-binary
capacity, is deferred to the appendix. The point stands: \textbf{politeness under
age-limited sensing is worth exactly what the neighbour's load is predictable
across the sensing delay, and a controller should measure that predictability and
adapt.}
