% nats-bursting -- INFOCOM Evaluation (Section 8). Drop-in; \input after the model.
% Preamble: \usepackage{booktabs} (optional). References sec:model, sec:controller,
% sec:thermal, sec:fabric, prop:bound from infocom_model.tex.
%
% HONESTY MARKERS: paragraphs tagged %PENDING% describe experiments not yet run;
% do not present them as results. Numbers below are measured (5-rep mean +/- 95% CI
% unless noted); see benchmarks/infocom/RESULTS.md and out/ms/.

\section{Evaluation}
\label{sec:eval}

We evaluate nats-bursting on a production testbed: a home node and an Atlas
workstation-class node running \texttt{nats-server} with JetStream enabled,
connected over three paths---co-located (loopback), Tailscale (mesh VPN), and a
single PAT-forwarded port (\texttt{:4222}). Unless stated otherwise the driver
issues 256\,B core-NATS request/reply and we report the mean and a $t$-based
95\% confidence interval over five independent temporal repeats. The study
answers five questions:
\textbf{(Q1)} how the federated control path shapes latency and throughput;
\textbf{(Q2)} what JetStream durability costs;
\textbf{(Q3)} how the control plane scales with the in-flight window;
\textbf{(Q4)} the prober detection delay $D$ that governs the bound of
Prop.~\ref{prop:bound}; and
\textbf{(Q5)} whether the politeness controller dominates naive and static
baselines on the goodput--compliance frontier.

\subsection{Control-plane latency and throughput (Q1)}
\label{sec:eval-rtt}
Co-located, request/reply RTT is $0.84$\,ms (p50) at $6855$\,msg/s. Over
Tailscale---one WAN hop to the same responder---RTT rises to $76\pm13$\,ms (p50),
$171\pm96$ (p95), $302\pm170$ (p99), and closed-loop throughput at a 16-message
window falls to $188\pm28$\,msg/s. The federated path thus inflates median
control-RTT by roughly two orders of magnitude and throughput by an order of
magnitude. Crucially the cost is \emph{round-trip latency}, not server work:
a synchronous, per-message control loop would be RTT-bound. This is the direct
empirical motivation for the \emph{epoch-paced} admission controller of
\S\ref{sec:controller} rather than per-message acknowledgement, and for the
wide-tail variance ($95\%$ CI on p99 spans $\pm170$\,ms) that any WAN-side
controller must tolerate.
%PENDING% The single-port \texttt{:4222} traversal tax (E2) is not yet measured;
%PENDING% it requires a PAT-forwarding window unavailable during these runs.

\subsection{Durability cost (Q2)}
\label{sec:eval-durability}
We compare JetStream publish-to-ack latency for file-backed (durable) and
memory (non-durable) streams over Tailscale. The medians are
$81.4\pm10.6$\,ms (durable) and $84.0\pm6.5$\,ms (non-durable), with overlapping
confidence intervals at every percentile. Over this WAN, file persistence is
therefore \emph{statistically indistinguishable} from in-memory: the round-trip
dominates the on-disk write, so durability is effectively free on the latency
axis. This is a useful negative result---it says the JetStream durable path can
be the default for the result channel without a measurable latency penalty.
%PENDING% A server-restart durability test (where memory should not survive and
%PENDING% file should) is deferred to a controlled maintenance window; the present
%PENDING% recovery check exercises only client reconnection.

\subsection{Control-plane scaling (Q3)}
\label{sec:eval-scaling}
Sweeping the in-flight window $w\in\{1,4,8,16,32\}$, throughput grows
sub-linearly: $11, 39, 83, 158, 283$\,msg/s respectively (95\% CIs widening from
$\pm2$ to $\pm128$). The curve is consistent with an RTT- and contention-bound
closed loop and with the AIMD window dynamics of \S\ref{sec:controller}: useful
in-flight work scales with the window until queueing and shared-link contention
flatten the gain---exactly the regime in which additive-increase probing and
multiplicative back-off are the appropriate control law.

\subsection{Subject interest (Q1, federation)}
\label{sec:eval-interest}
On a single server, publishing to a subscribed subject yields $300/300$ delivery
with zero cross-subject leakage and propagation p50 $169$\,ms, confirming subject
isolation and correct interest routing.
%PENDING% The asymmetric leaf-federation interest behavior central to
%PENDING% \S\ref{sec:fabric} (and the result-webhook return path it motivates)
%PENDING% requires a two-leaf topology and is not yet measured; the single-server
%PENDING% result above is a routing sanity check, not the federation claim.

\subsection{Detection delay and the violation bound (Q4)}
\label{sec:eval-detection}
$D$ has two measured components. The prober's per-call cost (read-only
\texttt{nvidia-smi}) is $27$--$33$\,ms (p50). The \emph{end-to-end} detection delay
---from launching a real GPU load until it is visible as utilization above the
$40\%$ floor (10 reps on a GV100, thermally guarded)---is \textbf{$2.46$\,s}
(mean; p50 $2.48$, range $2.29$--$2.61$), dominated by CUDA cold-start rather than
the poll. The relevant $D$ depends on the question: detecting a \emph{newly
admitted} pod's (un)productivity incurs the $\sim\!2.5$\,s cold-start term, while
probing \emph{already-running} pods costs only the $\sim\!27$\,ms poll.
Substituting into Prop.~\ref{prop:bound}, $\rho\approx D\alpha/((1-\beta)\bar a)$,
the $2.46$\,s term makes $D$ a \emph{first-order} contributor to the achievable
compliance---not a rounding error---which is precisely the systems$\leftrightarrow$theory
hinge the bound is meant to expose: a guest that admits aggressively (large
$\alpha$) pays for it in over-admission scaled by this measured cold-start delay.

\subsection{End-to-end goodput versus compliance (Q5)}
\label{sec:eval-e2e}
%PENDING% THIS IS THE HEADLINE EXPERIMENT AND IS NOT YET RUN. Do not present as a result.
\emph{(Design; results pending.)} The end-to-end experiment submits a burst of
$B$ jobs under three policies---naive (submit-all), static-rate, and the AIMD
admission controller---into an NRP namespace within fair-use
(\S\ref{sec:model}: pod cap $K{=}4$, GPU-utilization floor $40\%$). We measure
burst-completion time, sustained GPU utilization, and the realized over-admission
rate $\rho$, and we compare the realized $\rho$ against the
Prop.~\ref{prop:bound} prediction parameterized by the $D$ measured in
\S\ref{sec:eval-detection}, closing the model$\leftrightarrow$measurement loop.
The hypothesis is that AIMD dominates both baselines on the goodput--compliance
frontier: naive violates the cap and the utilization floor under contention;
static under-utilizes when capacity is available. \textbf{Until this experiment
is run against a contended, multi-tenant cluster, the central claim of the
paper---that polite bursting maximizes goodput while bounding policy
violation---remains unvalidated.}

\subsection{Summary and threats to validity}
\label{sec:eval-threats}
The measurements above characterize the \emph{federated control substrate} with
confidence intervals and even self-correct earlier single-run artifacts (a
spurious low throughput sample and an apparent durability tail gap both vanished
under repetition). They do not yet demonstrate the controller in action.
Principal threats: (i) a single home$\leftrightarrow$node path with \emph{no
competing tenants}, so the contended capacity $a_t$ of \S\ref{sec:model} is held
artificially constant; (ii) federation claims measured on one server, not a leaf
topology; (iii) the AIMD and Kalman controllers (\S\ref{sec:controller},
\S\ref{sec:thermal}) are analyzed but not yet measured under load; and (iv) no
comparison against existing bursting systems. \S\ref{sec:eval-e2e} and the leaf,
partition (E6), and full-$D$ experiments are required before the evaluation
supports the paper's thesis.
