\documentclass[conference]{IEEEtran}
\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning}
\usepackage[hidelinks]{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

% ---- Double-blind toggle -------------------------------------------------
% \blindtrue  => anonymized submission PDF (INFOCOM double-blind).
% \blindfalse => de-anonymized camera-ready (restores author + real names).
\newif\ifblind
\blindtrue
\newcommand{\sysname}{\ifblind PoliteBurst\else nats-bursting\fi}
\ifblind\hypersetup{pdfauthor={},pdftitle={},pdfsubject={},pdfkeywords={},pdfcreator={}}\fi
% -------------------------------------------------------------------------

\begin{document}

\title{Polite Bursting: Feedback Admission Control under
Age-Limited Sensing on Shared Clusters}

\author{\IEEEauthorblockN{\ifblind Anonymous Author(s)\else Andrew H. Bond\fi}
\IEEEauthorblockA{\ifblind Paper submitted for double-blind review\else San Jose State University\\
andrew.bond@sjsu.edu\fi}}

\maketitle

\begin{abstract}
Research and AI teams burst GPU/ML batch workloads onto shared, multi-tenant
clusters where they are \emph{guests}: bound by fair-use policy (pod caps,
utilization floors) and forced to sense a neighbour-driven, unobservable capacity
through a \emph{stale} probe. Naive submission violates policy; static
rate-limiting wastes throughput; and adaptive control sometimes helps and sometimes
does not. We explain \emph{when} by casting polite bursting as chance-constrained
admission under age-of-information: the controller acts on a capacity estimate of
age $D$. Our central result is an exact \emph{feedback-usefulness limit}: for a
symmetric two-state capacity, the goodput advantage of any causal admission
controller over the best static budget equals \emph{half the autocorrelation of
available capacity at the sensing lag}, $\Delta(D)=\tfrac12 r(D)$---a value that
stays \emph{achievable} for any symmetric stationary capacity (via a memoryless
threshold), with the general converse governed by the capacity's $\beta$-mixing
rate. It yields a sharp impossibility (no controller beats
static once $D$ exceeds the neighbour's coherence time $\tau_c$), a matching
achievability, and a \emph{regime-adaptive} controller that estimates $\tau_c$
online and switches between closed- and open-loop admission to track the limit
uniformly. We validate the law in simulation---the optimal policy sits on
$\tfrac12 r(D)$ to within $0.013$, practical AIMD degrades below static exactly when
$\tau_c\!\lesssim\!D$, and the regime-adaptive controller never falls below
static (a live GPU run confirms its online $r(D)$ estimate lands on the curve and
the switch's politeness dividend grows as feedback fails)---and on a workstation
testbed under \emph{real, manufactured GPU contention}
(with the full submission path verified end-to-end on NRP Nautilus): under structured
contention feedback admission is the polite-efficient knee, $1.41\times$ static's
goodput at $4\times$ lower over-admission \emph{than naive} while sparing the
competing tenant a greedy policy harms, and up to $3.2\times$ under a self-imposed
budget. The pre-registered
memoryless regime where it breaks even is exactly the $r(D)\!\to\!0$
prediction---the regime the adaptive controller is built to detect and sidestep. The
system and a reproducible artifact are
\ifblind available to reviewers\else open source\fi.
\end{abstract}

\section{Introduction}
\label{sec:intro}
A growing fraction of machine-learning work is \emph{bursty}: long stretches of
local development punctuated by short, intense bursts of batch jobs (sweeps,
fine-tunes, evaluations) that briefly need far more GPUs than a team owns.
Shared, opportunistic clusters such as the National Research Platform (NRP)
\cite{nautilus} make that capacity available, but on strict \emph{guest} terms:
fair-use policy caps concurrent pods, demands sustained utilization of what you
request, bans sleep-only or idle workloads, and reclaims resources on completion.
The guest is also \emph{networked} as an outsider---often reachable only through a
single forwarded port or a mesh VPN, with no inbound cluster access.

In this setting the two obvious strategies both fail. \emph{Naive} submission
fires all work at once: it grabs more than its fair share, leaves pods pending or
under-utilized, and trips the very thresholds that get a namespace banned.
\emph{Static} rate-limiting is safe but leaves the budget idle whenever capacity
is available, wasting the guest's own throughput. Neither adapts.

We argue that being a good guest is a \emph{control} problem whose answer is set by
\emph{how stale the guest's view of the cluster is}, and we build that thesis into
\sysname, a container-native bursting fabric. Our contributions:
\begin{itemize}
\item An \textbf{age-of-information formulation} of polite bursting
(\S\ref{sec:model}) and a \textbf{feedback-usefulness limit}
(\S\ref{sec:limit}): with a capacity estimate of age $D$, the goodput advantage of
\emph{any} causal admission controller over the best static budget is
$\Delta(D)=\tfrac12 r(D)$, half the autocorrelation of available capacity at the
sensing lag---exactly so for a symmetric two-state capacity (Thm.~\ref{thm:limit}),
and achievably in general, with the converse governed by $\beta$-mixing
(App.~\ref{app:beta})---an impossibility once $D$ exceeds the
neighbour's coherence time $\tau_c$ (Cor.~\ref{cor:imp}), lifted to any pod cap by a
layer-cake reduction (Prop.~\ref{prop:layercake}).
\item A \textbf{regime-adaptive controller} (\S\ref{sec:controller}) that estimates
$\tau_c$ online from the probe stream and switches between closed- and open-loop
admission, provably tracking the limit uniformly in $\tau_c$
(Thm.~\ref{thm:adapt}) with a deterministic hard-cap safety guarantee
(Prop.~\ref{prop:safety}).
\item A \textbf{realization} (\S\ref{sec:fabric}) in which the controller runs
in-cluster and senses/actuates across a single-port WAN leaf link that fixes the
sensing age $D$; the fabric's own design and characterization are the subject of a
companion systems paper~\cite{companion} and are out of scope here.
\item An \textbf{evaluation} (\S\ref{sec:eval}) that validates the law in
simulation (optimal policy on $\tfrac12 r(D)$ within $0.013$; adaptive control
never below static) and, on a workstation testbed under \emph{real GPU contention}
(full submission path verified end-to-end on NRP Nautilus), shows feedback admission
is the polite-efficient knee that spares the competing tenant---while the
pre-registered memoryless regime where it breaks even is exactly the
$r(D)\!\to\!0$ prediction the adaptive controller detects and sidesteps.
\end{itemize}

\begin{figure}[t]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[
  font=\footnotesize, node distance=5mm,
  box/.style={draw, rounded corners, align=center, minimum height=7mm, inner sep=2pt},
  >={Stealth[]}]
\node[box] (client) {home node\\client + driver};
\node[box, below=of client] (hub) {NATS hub};
\node[box, right=14mm of hub] (leaf) {NATS leaf\\(in-cluster)};
\node[box, above=of leaf] (ctrl) {controller\\(admission)};
\node[box, right=10mm of ctrl] (pods) {GPU worker\\pods};
\draw[<->] (client) -- (hub) node[midway,right]{burst.submit};
\draw[<->] (hub) -- (leaf) node[midway,above]{leaf link};
\draw[->] (leaf) -- (ctrl);
\draw[->] (ctrl) -- (pods) node[midway,above]{Job};
\draw[->] (pods) to[bend right=25] node[midway,below]{results.$\langle$id$\rangle$} (leaf);
\draw[->,dashed] (pods.north) to[bend right=30]
  node[midway,above,align=center]{congestion $s_t$\\[-1pt](age $D$)} (ctrl.north);
\end{tikzpicture}}
\caption{\sysname{} architecture. The home client submits over a federated
NATS leaf link; the in-cluster controller admits jobs (pacing concurrency to the
politeness budget) and launches GPU worker pods, whose results route back through
the leaf. The dashed arc is the stale feedback the controller acts on: an age-$D$
congestion signal $s_t$ whose delay bounds the achievable advantage
(Thm.~\ref{thm:limit}).}
\label{fig:arch}
\end{figure}

\section{Background and Motivation}
\label{sec:bg}
\textbf{Host policy.} The reference host (NRP Nautilus \cite{nautilus}) publishes a
fair-use policy that our design treats as ground truth: at most $K{=}4$ concurrent
pods per user; per running pod, utilization relative to \emph{request} must stay
in published bands (GPU $\ge 40\%$, CPU $20$--$200\%$, RAM $20$--$150\%$); jobs
must do real work (sleep-only jobs are banned) and release resources on
completion; idle storage volumes are reclaimed. Admins monitor utilization and
ban namespaces that hold under-utilized requests---so politeness is not optional.

\textbf{Network reality.} The guest reaches the cluster over a single
PAT (port address translation)-forwarded port or a mesh VPN, with no inbound access. A control plane must
therefore live behind that boundary and communicate over a messaging fabric that
traverses it---motivating the NATS leaf-node design of \S\ref{sec:fabric}.

\textbf{Why existing tools are not enough.} Workflow and function-execution
frameworks---HTCondor \cite{htcondor}, funcX/Globus Compute \cite{funcx}, Parsl
\cite{parsl}, Ray \cite{ray}, Dask \cite{dask}---scale workers to demand but carry
no notion of a host \emph{politeness budget}; left to default scaling they
over-provision under contention. We instead make self-restraint a first-class,
controlled objective.

\input{infocom_aoi_model}

\section{Realization: Sensing over a Federated Fabric}
\label{sec:fabric}
The controller runs \emph{in-cluster}, behind the network boundary, and reaches the
home node over a NATS~\cite{nats} \emph{leaf} link that needs only one outbound port
(Fig.~\ref{fig:arch}): it probes cluster occupancy and actuates admission across that
link, submissions flowing on \texttt{burst.submit} and results returning on
\texttt{results.>} by interest propagation, with no inbound cluster access. What
matters for the controller is that this view is \emph{aged}---the leaf round-trip and
the CUDA cold-start of a freshly admitted pod together set the sensing age $D$ of
\S\ref{sec:limit}, which we measure in \S\ref{sec:eval}. The tensors crossing the link are
compressed by random-rotation quantization \cite{turboquantpro,turboquant}; this aids
bandwidth on the hop, but the sensing age $D$ is dominated by CUDA cold-start, not
transport (\S\ref{sec:eval}), so compression does not reduce $D$. The fabric's own
design---leaf-federation tradeoffs, JetStream durability, single-port traversal---and
the transport and node-level thermal subsystems are a systems contribution developed
and characterized in a companion paper~\cite{companion}; here the fabric is simply the
medium that makes capacity sensing possible, and stale.

\section{Implementation}
\label{sec:impl}
\sysname{} is a Go control plane (decider, prober, submitter, NATS bridge,
back-off modules) plus a Python client and an IPython cell magic
\cite{natsbursting}. The controller runs \emph{in-cluster} under a minimal
namespace-scoped ServiceAccount (Job create/get/delete; pod read), connects to the
in-cluster NATS leaf, and renders each \texttt{JobDescriptor} into a Kubernetes
Job; node-targeting (a \texttt{nodeSelector} on the descriptor) pins guest pods to
non-reserved GPU types. We verified the full path live on NRP Nautilus
\cite{nautilus}: a submit from the home node propagated hub$\to$leaf$\to$controller
and launched an A10 GPU worker that ran real CUDA matmul and published its result
back through the leaf to the home driver. The system, the controller/worker
container images, and the transport/right-sizing packages
\cite{turboquantpro,batchprobe} are \ifblind available as open-source artifacts (repositories anonymized for review)\else open source\fi.

\input{infocom_eval}

\input{crossdomain_predictive}

\section{Related Work}
\label{sec:related}
\textbf{Bursting and function execution.} HTCondor \cite{htcondor}, funcX/Globus
Compute \cite{funcx}, Parsl \cite{parsl}, Ray \cite{ray}, and Dask \cite{dask}
provide elastic task execution and provider-driven scaling, but treat the host as
owned capacity to fill; none model a fair-use politeness budget or admission under
age-limited sensing. They are complementary substrates---a head-to-head
dispatch/throughput comparison against them, and the fabric that carries our control
plane, is reported in the companion systems paper~\cite{companion}. \textbf{Messaging.}
The control plane rides NATS \cite{nats} leaf federation rather than a bespoke RPC
layer.
\textbf{Congestion and scavenger control.} Our controller is AIMD in the lineage of
additive-increase/multiplicative-decrease congestion avoidance \cite{chiujain}, and
its \emph{polite} objective---yield to incumbents, use only slack---aligns with
low-priority/scavenger transports such as LEDBAT \cite{ledbat} and TCP-Nice
\cite{tcpnice}. Those protocols are engineered to back off under one-sided loss or
delay signals; we differ in two ways. First, the objective is \emph{two-sided}
(stay under a pod cap \emph{and} above a utilization floor), which admits an
over-admission metric rather than pure loss. Second, and more fundamentally, we do
not merely propose a back-off rule: we prove a \emph{limit} on what any such rule
can achieve. \textbf{Age of information.} That limit is an information-freshness
statement---the value of feedback is set by the age $D$ of the capacity estimate
relative to the neighbour's coherence time---placing polite bursting in the
age-of-information line of work \cite{aoi} but on the \emph{control} rather than the
sampling side: we bound the achievable goodput--politeness frontier as a function of
$D$ and give a controller that adapts to the measured regime. To our knowledge, no
prior work frames shared-cluster bursting as admission control under age-limited
sensing, proves a feedback-usefulness limit for it, or ties an over-admission bound
to a measured detection delay.

\section{Conclusion}
\label{sec:concl}
Being a good guest on a shared cluster is a control problem, and how well it can be
solved is fixed by how stale the guest's view of the cluster is. \sysname{} makes
that precise: an age-of-information formulation, a feedback-usefulness limit
$\Delta(D)=\tfrac12 r(D)$ that says closed-loop admission is worth exactly the
autocorrelation of capacity at the sensing lag, and a regime-adaptive controller
that measures its own predictability and tracks the limit. The limit is validated
at the controller level (the optimal policy sits on $\tfrac12 r(D)$ to within
$0.013$; AIMD falls below static precisely when $\tau_c\!\lesssim\!D$) and on a
workstation testbed under real GPU contention (the full submission path verified
end-to-end on NRP Nautilus): under structured contention feedback admission is the
polite-efficient knee---$1.41\times$ static's goodput at a quarter of naive's
over-admission while sparing the competing tenant a greedy policy harms---and up to
$3.2\times$ under a self-imposed budget. The one regime where feedback breaks even,
under memoryless contention that outpaces detection, is not a disappointment but the
theory's own prediction ($r(D)\!\to\!0$, Cor.~\ref{cor:imp})---and the regime the
adaptive controller detects and sidesteps. Open problems remain: a production multi-tenant trace (ours is
manufactured contention at $C{=}2$), extending the exact limit beyond the binary
layer to general capacity processes via the $\beta$-mixing lower bound, and
predictive controllers that estimate $r(D)$ rather than only its regime.

\appendices
\input{infocom_appendix}

\bibliographystyle{IEEEtran}
\bibliography{infocom}

\end{document}
