% ============================================================================
% ORPHANED / NOT COMPILED: infocom.tex does NOT \input this file. The live model
% is infocom_aoi_model.tex; the Kalman/PID thermal micro-tier here moves to the
% CANOPIE-HPC companion. Do NOT \input alongside infocom_aoi_model.tex -- it would
% duplicate labels (sec:model, sec:controller, prop:safety, prop:bound) and reuse
% q, r for Kalman noise variances (colliding with q(D), r(D)). Kept for reference.
% ============================================================================
% nats-bursting -- INFOCOM model sections (drop-in).
% Preamble (add to main paper if absent):
%   \usepackage{amsmath,amssymb}
%   \newtheorem{proposition}{Proposition}

\section{System Model and Problem Formulation}
\label{sec:model}

A \emph{guest} bursts batch GPU/ML \emph{pods} onto a shared, multi-tenant
cluster governed by fair-use policy. Time is slotted into control epochs
$t=0,1,2,\dots$. Let $w_t\in\{0,\dots,K\}$ be the guest's concurrency window
(in-flight pods), where $K$ is the policy \emph{pod cap}. Let $a_t$ be the number
of pods the cluster can \emph{productively} run for the guest at epoch $t$; $a_t$
is set by other tenants, time-varying, and \emph{not directly observable}. The
useful goodput is $G_t=\min(w_t,a_t)$, while $(w_t-a_t)^{+}$ pods are
\emph{over-admitted}: they sit pending or idle and pull the guest's sustained
utilization below the policy floor $U$, marking it a poor citizen. A binary
\emph{congestion signal} $s_t$ fires when over-admission is detected (a pod
pending beyond $\tau$, or measured utilization below $U$), with a
\emph{detection delay} $D$ epochs induced by the prober.

\paragraph{Host policy (concrete instance).}
The reference host (NRP Nautilus) publishes a fair-use policy our formulation
encodes directly: at most \emph{$K=4$} concurrent pods per user; and, for every
\emph{running} pod, utilization relative to its \emph{request} must stay within
published bands---GPU $\ge 40\%$, CPU $\in[20\%,200\%]$, RAM $\in[20\%,150\%]$;
pods must perform real work (jobs that merely \texttt{sleep} are banned) and
release their resources on completion. Two design arms map onto this: a request
right-sizer (\S\ref{sec:thermal}; \ifblind our right-sizing tool\else batch-probe\fi) keeps CPU/RAM/GPU \emph{within} the
bands by not over-requesting, while the admission controller (\S\ref{sec:controller})
keeps admitted pods \emph{above} the $40\%$ GPU floor by not over-admitting; clean
exit plus TTL frees the request. Accordingly we take $U=0.4$ (the GPU floor) and
$K=4$, and treat a pod that falls below $U$ as over-admitted.

\paragraph{Problem (Polite Bursting).}
Let $\rho$ denote the long-run over-admission fraction,
\begin{equation}
\rho \triangleq \limsup_{T\to\infty}\tfrac1T\textstyle\sum_{t<T}\mathbf 1[w_t>a_t].
\end{equation}
The guest then solves
\begin{equation}
\max_{\{w_t\}}\ \sum_t \min(w_t,a_t)\ \ \text{s.t.}\ \
\underbrace{w_t\le K\ \forall t}_{\text{hard}},\ \
\underbrace{\rho \le \epsilon}_{\text{policy}}.
\end{equation}
The constraint is two-sided---stay \emph{below} the cap and \emph{above} the
utilization floor---distinguishing it from one-sided (loss-only) congestion control.

\section{The Politeness Controller}
\label{sec:controller}
Politeness is enforced by control at two scales: a \emph{macro} admission loop
over the federated messaging fabric, and a \emph{micro} thermal loop at the edge node.

\subsection{Macro tier: admission control (AIMD)}
The guest adapts its window by additive-increase / multiplicative-decrease:
\begin{equation}
w_{t+1}=
\begin{cases}
\min(w_t+\alpha,\,K), & s_t=0\ \text{(clean)},\\[2pt]
\max(\lceil\beta w_t\rceil,\,w_{\min}), & s_t=1\ \text{(congested)},
\end{cases}
\end{equation}
with additive step $\alpha>0$ and back-off factor $\beta\in(0,1)$.
We assume \textbf{(A1)} stationary capacity $a_t\equiv a$, $w_{\min}\le a\le K$;
\textbf{(A2)} reliable detection within $D$ epochs of an overshoot; \textbf{(A3)}
fixed $\alpha,\beta,w_{\min},K$.

\begin{proposition}[Hard-cap safety]\label{prop:safety}
Under the clamped update, $w_t\le K$ for all $t$ deterministically; the pod-cap
policy is never violated.
\end{proposition}
\begin{proof}
Additive increase is clamped to $K$ and multiplicative decrease cannot increase
$w_t$; the claim follows by induction on $t$.
\end{proof}

\begin{proposition}[Steady-state over-admission bound]\label{prop:bound}
Under \textnormal{(A1)--(A3)} the window forms a periodic sawtooth about $a$, and
the long-run over-admission fraction satisfies
\begin{equation}
\rho \;=\; \frac{D\,\alpha}{(1-\beta)\,(a+D\alpha)}
\;\approx\; \frac{D\,\alpha}{(1-\beta)\,a}.
\end{equation}
Hence for any target $\epsilon\in(0,1)$, choosing $\alpha\le \epsilon(1-\beta)a/D$
guarantees $\rho\le\epsilon$; as $D\to0$, $\rho\to0$.
\end{proposition}
\begin{proof}[Proof sketch]
After a backoff the window restarts at $\beta(a+D\alpha)$ and rises by $\alpha$
per clean epoch, reaching $a$ after $\big(a-\beta(a+D\alpha)\big)/\alpha$ epochs
(no violation), then spends $D$ further epochs above $a$ before detection triggers
a backoff at peak $a+D\alpha$. Each cycle has $\approx D$ violating epochs out of
length $L=a(1-\beta)/\alpha + D(1-\beta)$; the ratio gives $\rho$.
\end{proof}

\paragraph{Goodput--compliance tradeoff.}
Mean window (hence goodput) rises as $\beta\to1$ while $\rho$ falls as
$\alpha\to0$; both slow convergence. The pair $(\alpha,\beta)$ thus traces a
convergence/goodput/compliance Pareto frontier, characterized in
\S\ref{sec:eval}. Under non-stationary capacity with bounded variation
$|a_{t+1}-a_t|\le\delta$, the bound loosens by an additive $O(\delta/\alpha)$
tracking term.

\subsection{Micro tier: Kalman-filtered thermal control}
\label{sec:thermal}
At the edge/home node, polite use of shared hardware means holding a thermal
setpoint while maximizing compute. We model node temperature $T_t$ under load
$n_t$ (threads / active SMs) as a first-order thermal system with Newtonian
cooling and process/measurement noise:
\begin{align}
T_{t+1} &= T_t + b\,n_t - c\,(T_t - T_{\mathrm{amb}}) + \eta_t, & \eta_t&\sim\mathcal N(0,q),\\
y_t &= T_t + \nu_t, & \nu_t&\sim\mathcal N(0,r).
\end{align}
A Kalman filter maintains $\hat T_t$ via the standard predict/update recursion,
suppressing sensor noise. The controller applies PID action with feedforward
\emph{lookahead}: using the model to predict $\hat T_{t+h}$ over horizon $h$, it sets
\begin{equation}
\begin{split}
n_t = \mathrm{clip}\big(n_{\max} &- K_p e_t - K_i\textstyle\sum_{k\le t} e_k \\
                       &- K_d\,\Delta e_t - K_{\!f}\,(\hat T_{t+h}-T^{\star})\big),
\end{split}
\end{equation}
with tracking error $e_t=\hat T_t-T^{\star}$, throttling \emph{before} the
setpoint $T^{\star}$ is breached. With a stable plant
($0<c<1$) and standard PID gain conditions the closed loop is input-to-state
stable about $T^{\star}$, and the feedforward term curbs overshoot under load
steps. This is the node-level analogue of the macro admission loop: both maximize
useful work subject to a shared-resource ceiling.
